---
title: "Create animated maps with stegomyia indices"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Create_animated_maps_with_stegomyia_indices}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

#  Make to maps Stegomyia Indices

## The origin of data

The "Programa Estatal de Vigilancia Entomología y Control Integral de Enfermedades Trasmitidas por Vector" carries out various activities for disease prevention, which are guided by surveillance activities that provide entomological risk indices, which is why a Package rStegomyia will be made to calculate the indices of entomological risk in an automatic and timely manner for decision making.

All activities are captured daily on a platform, where it is possible to download the data for reporting and analysis. In the case of Surveillance by Ovitrampa (VO), the platform calculates the risk indices but for the Entomological Study (EE) activity it calculates them per block worked, these indices are required per risk location on a weekly or daily basis.

It is necessary to be able to calculate epidemiological week (SE) risk indices by type of study, risk locality and by epidemiological week in an automated manner.

The EE activity evaluates a sample before (Encuesta) and after (Verificacion) of the comprehensive control activities to determine the entomological risk that exists in a large area to be worked on (Sectors or Locality) and if it requires reinforcing control and prevention actions.

```{r setup}
library(rStegomyia)
library(gganimate)
library(transformr)
library(gifski)
```
## Load raw data 

Download the file .txt of Entomology Study from plataform "Vigilancia Entomológica y Control Integral del Vector".
Select the variables for your analysis and load the data of the Entomology Study of file .txt  to make dataframe.
If the variables are not specified, the variables that are going to be selected by default are:  "Tipo de Estudio", "Jurisdiccion", "Localidad", "Sector", "Fecha de Inicio", "Semana Epidemiologica", "Casas Revisadas", "Casas Positivas", "Total de Recipientes con Agua", "Total de Recipientes Positivos".

```{r}
#path2016<- c("C:/Users/Javier Edgar Verdugo/Desktop/ELB/EE/2016/")
#path2017<- c("C:/Users/Javier Edgar Verdugo/Desktop/ELB/EE/2017/")
#path2018<- c("C:/Users/Javier Edgar Verdugo/Desktop/ELB/EE/2018/")
#path2019<- c("C:/Users/Javier Edgar Verdugo/Desktop/ELB/EE/2019/")
#path2020<- c("C:/Users/Javier Edgar Verdugo/Desktop/ELB/EE/2020/")
path2021<- c("C:/Users/Javier Edgar Verdugo/Desktop/ELB/EE/2021/")

pathalfa <- path2021

path0 <-paste0(pathalfa,"DescargaEntomologicoe26.txt")

path_raw_data <- path0

df_lrd <- load_raw_data(path_raw_data)

path1 <-paste0(pathalfa,"df_lrd.csv")

write_csv(df_lrd, file = path1)
 
head(df_lrd)
```
## Clean raw data

Next, clean and give format to the dataframe´s variables of the Entomology Study  

```{r}

 df_crd <- clean_raw_data(df_lrd,
               path_out = path1
               )

 head(df_crd)

```
## Calcul Stegomyia indices by type of study and geographic area  

Select study type and sectors to analize the dataframe of the Entomology Study. If the study type is not specified, "Verificacion" is going to be selected by default. The function will calculate Stegomyia Indices  and Index Status by Sector. Use that data to make a new dataframe
```{r}
path2 <-paste0(pathalfa,"statusindicesector.csv")


#df_crd <- filter(df_crd, Localidad == "HERMOSILLO")

 df_crd <- df_crd %>%
   filter(Localidad == "HERMOSILLO")

sectors <- unique(df_crd$Sector, incomparables = FALSE)
sectors <- factor(sectors)

df_sitsgis <- get_stegomyia_indices_by_type_of_study_and_geo_is(df_crd,
                                                                st = "Verificacion",
                                                                var= sectors,
                                                                path_out = path2
                                                                )
head(df_sitsgis)

```
## Create the maps with status of Stegomyia indices with ggplot

Use the new dataframe created to make maps for Stegomia Index Status with the sectors of Locality "Hermosillo"

```{r}

list_of_maps <-get_maps_stegomyia_indices(df_sitsgis)

list_of_maps
```
## Create the maps with status of Stegomyia indices with leaflet 

Use the new dataframe created to make maps for Stegomia Index Status with the sectors of Locality "Hermosillo"

```{r}

Stegomyia_indx_map <-get_1map_stegomyia_indices(df_sitsgis)

Stegomyia_indx_map
```
## Create the maps with status of Stegomyia indices with gganimate 

### Sellec infomation

prepare the data so that you select by type of study, month worked and save it for each of the Stegomyia indices
```{r}
output_dir <-paste0(pathalfa,"maps/")

tipos_de_estudio <- c("Verificacion",
                      "Encuesta")
nombres_indices <- c("index_status_HI",
                     "index_status_CI",
                     "index_status_BI")

#df_crd <- dplyr::filter(df_crd,
#                        Localidad == "HERMOSILLO")
df_mes <- df_crd %>%
  dplyr::select("Fecha_de_Inicio")

list_mes <- unique(lubridate::month(df_mes$Fecha_de_Inicio))

```

### Save all maps

create a map by type of study, month worked and save it for each of the Stegomyia indices. Also save an image of each map created

```{r}

for (mes in list_mes) {
  for (tipo_estudio in tipos_de_estudio) {
      df_mes  <- df_crd %>%
      filter(lubridate::month(Fecha_de_Inicio) == mes)
      
      try(df_sitsgis <- get_stegomyia_indices_by_type_of_study_and_geo_is(
        df_mes,
        st = tipo_estudio, 
       var = df_mes$Sector, 
       path_out = file.path(paste0(output_dir,"statusindices_",
                                  tipo_estudio,
                                  "_mes_",
                                  mes,
                                  ".csv"))
        )
      )
      try(pathtem <-paste0(output_dir,
                       "statusindices_",
                       tipo_estudio,
                       "_mes_",
                       mes,
                       ".csv"))
      try(dftem <- read_csv(pathtem))
      try(dftem <- dftem %>%
        mutate(Mes = mes))
      try(write_csv(dftem, pathtem))
    
    try(list_of_maps <- get_maps_stegomyia_indices(df_sitsgis))
    
      try(for (i in seq_along(list_of_maps)) {
        nombre_archivo <- file.path(output_dir, 
          paste0("mapa_",
                 tipo_estudio,
                 "_mes_",
                 mes,
                 "_",
                 nombres_indices[i],
                 ".png")
        )
        png(filename = nombre_archivo)
        print(list_of_maps[[i]])
        dev.off()
    }
    )
  }
}
```
create a database that contains the information of the Sectors worked (such as "Encuesta") for month and the Stegomyia indices

```{r}
archivos <- list.files(path = output_dir, pattern = "^statusindices_Verificacion_mes_", full.names = TRUE)

df_totalv <- purrr::map_df(archivos, read.csv)

write.csv(df_totalv, file = file.path(output_dir, "total_statusindices_Verificacion.csv"), row.names = FALSE)
```

### Crear dataframes for "Encuesta" and "Verificacion"

create a database that contains the information of the Sectors worked (such as "Verificacion") for month and the Stegomyia indices

```{r}
archivos <- list.files(path = output_dir, pattern = "^statusindices_Encuesta_mes_", full.names = TRUE)

df_totale <- purrr::map_df(archivos, read.csv)

write.csv(df_totale, file = file.path(output_dir, "total_statusindices_Encuesta.csv"), row.names = FALSE)
```

Select the created study type dataframes and join with the shape files of Sectors of Hermosillo to create animated maps

```{r}
m1 <- shm1
m0 <- shm2
df_totale <- df_totale %>%
    mutate(Sector = as.factor(Sector))

w2e <- st_transform(m1,
                     crs =4326)
  w2e <- w2e %>%
    rename(Sector = SECCION)%>%
    mutate(Sector = as.factor(Sector))
  w2e_ <- w2e %>%
    left_join(df_totale,
                         by = "Sector")
  w2e_ <- w2e_ %>% drop_na()
  
  w2e_ <- st_transform(w2e_,
                     4326)

df_totalv <- df_totalv %>%
    mutate(Sector = as.factor(Sector))
  
w2v <- st_transform(m1,
                     crs =4326)
  w2v <- w2v %>%
    rename(Sector = SECCION)%>%
    mutate(Sector = as.factor(Sector))
  
  w2v_ <- w2v %>%
    left_join(df_totalv,
                         by = "Sector")
  w2v_ <- w2v_ %>% drop_na()
  
  w2v_ <- st_transform(w2v_,
                     4326)  
m0 <- st_transform(m0,
                     crs = 4326
                     )

 colores_id <- c(
    "Optimo" = "blue",
    "Bueno" = "green",
    "Alarma" = "yellow",
    "Emergencia" = "red"
  )
```
## Create animated maps

Use the new dataframe created to make animate maps for Stegomia Index Status with the sectors of Locality "Hermosillo" of type of study in "Encuesta"

```{r}
pe_HI <- ggplot() +
    geom_sf(data = m0,
            fill = NA,
            size = 0.9) +
    geom_sf(data = w2e_,
            aes(fill = index_status_HI),
            color = "black",
            size = 0.9) +
    coord_sf() +
    labs(title = "Stegomyia House Index in Hermosillo(Encueta)") +
    scale_fill_manual(values = colores_id) +
  theme_minimal()

p_animado <- pe_HI +
  transition_states(Mes,
                    transition_length = 20,
                    state_length = 1) +
  ease_aes('linear')+ 
  labs(title = "Stegomyia House Index in Hermosillo(Encueta)",
       subtitle = 'Mes:{closest_state}') +   
  theme(plot.subtitle = element_text(size = 10)) 

pe_HI_gif <- animate(p_animado,
                         nframes = 12,
                         fps = 2,
                         renderer = gifski_renderer(paste0(output_dir,"pe_HI_mapa.gif")))

pe_CI <- ggplot() +
    geom_sf(data = m0,
            fill = NA,
            size = 0.9) +
    geom_sf(data = w2e_,
            aes(fill = index_status_CI),
            color = "black",
            size = 0.9) +
    coord_sf() +
    labs(title = "Stegomyia Container Index in Hermosillo(Encueta)") +
    scale_fill_manual(values = colores_id) +
  theme_minimal()

p_animado <- pe_CI +
  transition_states(Mes,
                    transition_length = 20,
                    state_length = 1) +
  ease_aes('linear')+ 
  labs(title = "Stegomyia Container Index in Hermosillo(Encueta)",
       subtitle = 'Mes:{closest_state}') +   
  theme(plot.subtitle = element_text(size = 10)) 

pe_CI_gif <- animate(p_animado,
                         nframes = 12,
                         fps = 2,
                         renderer = gifski_renderer(paste0(output_dir,"pe_CI_mapa.gif")))

pe_BI <- ggplot() +
    geom_sf(data = m0,
            fill = NA,
            size = 0.9) +
    geom_sf(data = w2e_,
            aes(fill = index_status_BI),
            color = "black",
            size = 0.9) +
    coord_sf() +
    labs(title = "Stegomyia Breteau Index in Hermosillo(Encueta)") +
    scale_fill_manual(values = colores_id) +
  theme_minimal()

p_animado <- pe_BI +
  transition_states(Mes,
                    transition_length = 20,
                    state_length = 1) +
  ease_aes('linear')+ 
  labs(title = "Stegomyia Breteau Index in Hermosillo(Encueta)",
       subtitle = 'Mes:{closest_state}') +  
  theme(plot.subtitle = element_text(size = 10)) 

pe_BI_gif <- animate(p_animado,
                         nframes = 12,
                         fps = 2,
                         renderer = gifski_renderer(paste0(output_dir,"pe_BI_mapa.gif")))

pe_HI_gif
```
```{r}
pe_CI_gif
```
```{r}
pe_BI_gif
```
Use the new dataframe created to make animate maps for Stegomia Index Status with the sectors of Locality "Hermosillo" of type of study in "Verificacion"

```{r}
pv_HI <- ggplot() +
    geom_sf(data = m0,
            fill = NA,
            size = 0.9) +
    geom_sf(data = w2v_,
            aes(fill = index_status_HI),
            color = "black",
            size = 0.9) +
    coord_sf() +
    labs(title = "Stegomyia House Index in Hermosillo(Verificacion)") +
    scale_fill_manual(values = colores_id) +
  theme_minimal()

p_animado <- pv_HI +
  transition_states(Mes,
                    transition_length = 20,
                    state_length = 1) +
  ease_aes('linear')+ 
  labs(title = "Stegomyia House Index in Hermosillo(Verificacion)",
       subtitle = 'Mes:{closest_state}') +   
  theme(plot.subtitle = element_text(size = 10))


pv_HI_gif <- animate(p_animado,
                         nframes = 12,
                         fps = 2,
                         renderer = gifski_renderer(paste0(output_dir,"pv_HI_mapa.gif")))

pv_CI <- ggplot() +
    geom_sf(data = m0,
            fill = NA,
            size = 0.9) +
    geom_sf(data = w2v_,
            aes(fill = index_status_CI),
            color = "black",
            size = 0.9) +
    coord_sf() +
    labs(title = "Stegomyia Container Index in Hermosillo(Verificacion)") +
    scale_fill_manual(values = colores_id) +
  theme_minimal()

p_animado <- pv_CI +
  transition_states(Mes,
                    transition_length = 20,
                    state_length = 1) +
  ease_aes('linear')+ 
  labs(title = "Stegomyia Container Index in Hermosillo(Verificacion)",
       subtitle = 'Mes:{closest_state}') +  
  theme(plot.subtitle = element_text(size = 10))

pv_CI_gif <- animate(p_animado,
                         nframes = 12,
                         fps = 2,
                         renderer = gifski_renderer(paste0(output_dir,"pv_CI_mapa.gif")))

pv_BI <- ggplot() +
    geom_sf(data = m0,
            fill = NA,
            size = 0.9) +
    geom_sf(data = w2v_,
            aes(fill = index_status_BI),
            color = "black",
            size = 0.9) +
    coord_sf() +
    labs(title = "Stegomyia Breteau Index in Hermosillo(Verificacion)") +
    scale_fill_manual(values = colores_id) +
  theme_minimal()

p_animado <- pv_BI +
  transition_states(Mes,
                    transition_length = 20,
                    state_length = 1) +
  ease_aes('linear')+ 
  labs(title = "Stegomyia Breteau Index in Hermosillo(Verificacion)",
       subtitle = 'Mes:{closest_state}') +  
  theme(plot.subtitle = element_text(size = 10))

pv_BI_gif <- animate(p_animado,
                         nframes = 12,
                         fps = 2,
                         renderer = gifski_renderer(paste0(output_dir,"pv_BI_mapa.gif")))

pv_HI_gif
```
```{r}
pv_CI_gif
```
```{r}
pv_BI_gif
```

